---
title: "Insect Abundance Mapping at Jasper Ridge Biological Preserve"
author: "Rachel Fox"
output: html_document
runtime: shiny
---

Data used for this project was collected in 2012 and 2013 under David Moreno-Mateos at Jasper Ridge Biological Preserve (Stanford University). All data was used with permission. GIS Shapefiles and weather data prior to April 2013 were provided by JRBP for expressed use in this project. Weather data from April 2013 and beyond was taken from Weather Underground weather station at JRBP, available online.  

```{r, echo=FALSE, warning=FALSE, message=FALSE}
suppressPackageStartupMessages(library(shiny))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(rgdal))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(rvest))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(maps))
suppressPackageStartupMessages(library(ggmap))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(require(gridExtra))

#Insect Counts and Timeseries
insectcount<- read.csv("./Data/insectcounts_na.csv", header=TRUE) %>%
              tbl_df() %>% 
              gather(family,n, 5:68) %>%
              mutate(date=mdy(date)) %>%
              transmute(date, bottle.name, location, trap.name, family, n)

#Timeseries- Weather
weather1<- read.csv("./Data/weatherdata.txt") %>% tbl_df() %>%
           mutate(date= ymd(Date))%>% 
          select(date, TemperatureAvgF)

weather <- read.csv("./Data/insect_weather.csv") %>% tbl_df %>%
            unite(Date, year,dayofyear, sep = "-") %>%
            transmute(date=parse_date_time(Date, "%y%j"),TemperatureAvgF=TemperatureAvgF) %>%
            bind_rows(weather1)
          

#Combining Insects with Trap Locations.
traplocations <- read.csv("./Data/traplocations.csv",header=TRUE) %>%
                 tbl_df()
locationcoord <- read.csv("./Data/locationzoom.csv", header=TRUE) %>%
                 tbl_df()

insects <- insectcount %>% 
                group_by(date, trap.name, family) %>%
                summarise(count=sum(n))
insectloc <- inner_join(insects, traplocations, by="trap.name") %>%
                tbl_df() 

zoomcoords <- insectcount %>%
              group_by(date, location, trap.name, family) %>%
              summarise(count=sum(n))%>%
              inner_join(traplocations, by="trap.name") %>%
              inner_join(locationcoord, by="location")


# for total abundance timeseries
insectsum <- insectcount %>% 
             group_by(date, family) %>%
             summarise(count=sum(n))
     
#Extracting data and creating data for plotting
##Trails

jr.trailsfile <- readOGR(dsn="./Data/JR_TRAILS", layer="JR_TRAILS", verbose=FALSE) %>%
                 spTransform(CRS("+proj=longlat +ellps=WGS84")) 

trails.map <- fortify(jr.trailsfile)
trail.data <-jr.trailsfile@data %>% tbl_df()

##Vegetation
jr.vegfile <- readOGR(dsn="./Data/JRVEG_SHAPEFILE", layer="JR_vegetation_2012", 
              verbose=FALSE) %>%
              spTransform(CRS("+proj=longlat +ellps=WGS84")) 

jr.vegfile@data$id = rownames(jr.vegfile@data)

veg.map <- fortify(jr.vegfile, region="id") %>% tbl_df()
veg.df <- inner_join(veg.map, jr.vegfile@data, by="id") %>%
          filter(CNDDB1 %in% c("Bog and Marsh", "Broad Leafed Upland Tree Dominated", "Coniferous Upland Forest and Woodland", "Grass and Herb Dominated Communities", "Riparian and Bottomland Habitat", "Scrub and Chaparral")) %>%
          rename(Vegetation=CNDDB1)

#Overall Map
#google map access
google.map <-
  get_map(location = "Searsville Lake, San Mateo County, CA", maptype = "hybrid", zoom = 16, color = "color")


```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
inputPanel(
  dateRangeInput("daterange", "Date range:",
               start = "2012-05-30",
               end   = "2013-07-10",
               format = "MM-dd-yy",
               min = "2012-05-30",
               max = "2013-07-10"
               ),
  
  selectInput("family", label = "Insect Family",
               choices = c("total.count", "s.acalyptrata", "acalyptrata", "s.calyptrata", "calyptrata", "auchenorrhynca", "o.auchenorrhyncha", "polyphaga", "o.polyphaga", 
"s.apocrita", "apocrita", "o.apocrita", "p.apocrita", "formicidae", 
"vespidae", "anthrophila", "chironomidae", "o.chironomidae", 
"sciaridae", "mycetophilidae", "o.mycetophilidae", "cecydomidae", 
"empidoidea", "o.empidoidea", "platypezidae", "phoroidea", "o.phoroi", 
"sternorrhyca", "o.sternorrhyncha", "psocoptera", "odonata", 
"syrphoidia", "o.syrphoidea", "s.lepidoptera", "lepidoptera", 
"o.lepidoptera", "lepidoptera.2", "psychodomorpha", "neuroptera", 
"tipulomorpha", "o.tipulomorpha", "heteroptera", "o.heterop", 
"tingidae", "plannipenia", "o.plannipenia", "tricoptera", "o.tricoptera", 
"symphyta", "ceratopogonidae", "xylophagomorpha", "tabanomorpha", 
"thysanoptera", "scathophagidae", "ephemeridae", "asilidae", 
"culicidae", "isoptera", "simuliidae", "lonchopteridae", "plecoptera", 
"rophidoptera", "dolichopodidae"), selected = "total.count", multiple= TRUE
  ),

selectInput("zoom", label="Zoom to Traps", choices=c("all", "L","S","O","M","R"), selected= "all"),

submitButton("Submit")
)

renderPlot({
      # Inputs and Variables.    
      startdate <- input$daterange[1]
      enddate <- input$daterange[2]
      date.range <- interval(startdate, enddate)
      
      familynames <- input$family
      zoomfocus <- input$zoom
      
      zoomfocus <- c("L")
      
google2.map <- get_map(location = ifelse(zoomfocus="all", "Searsville Lake, San Mateo County, CA",
                                        ifelse(zoomfocus="L", "37.40498, -122.2388",
                                          ifelse(zoomfocus="S","37.39544, -122.2387",
                                                 ifelse(zoomfocus="O", "37.40572, -122.2369", 
                                                        ifelse(zoomfocus="M", "37.40167,-122.2423",
                                                               ifelse(zoomfocus="R", "37.41012, -122.2384",.)))))), 
                      maptype = "hybrid", 
                      zoom = ifelse(zoomfocus="all", 16,
                                        ifelse(zoomfocus="L", 18,
                                          ifelse(zoomfocus="S", 18,
                                                 ifelse(zoomfocus="O", 18,
                                                         ifelse(zoomfocus="M", 18,
                                                                 ifelse(zoomfocus="R", 18,.)))))), color = "color")

ggmap(google2.map)
                      
                      
      insectgraph <- filter(insectloc, family %in% familynames ) %>%
                     filter(date %within% date.range)
      
      insecttime <- filter(insectsum, family %in% familynames) %>%
                    filter(date %within% date.range)
      
      weather <- filter(weather, date %within% date.range)
    
      
      zoomgraphL <-
      zoomgraphM <-
      zoomgraphS <-
      zoomgraphO <-
  
      
# Make map.  
map <-  ggmap(google.map)+
        geom_polygon(data=veg.df, 
                     aes(x=long, y=lat, group=group, fill=Vegetation),alpha=0.3) +
        geom_path(data=trails.map, 
                     aes(x=long, y=lat, group=group),col="black", size=0.2) +
        geom_point(data=insectgraph, 
                     aes(x=long, y=lat, size=count, color=family), alpha=0.5)+
        scale_color_brewer(palette="Set1")+
        scale_size_continuous(range = c(1, 18))+
        coord_map() +
        xlab("Longitude") + ylab("Latitude")+
        theme_bw()
  
#Timeseries
i.timeseries <- ggplot(insecttime, aes(x=date, y=count, col= family))+
                geom_line()+
                scale_color_brewer(palette="Set1")+
                xlab("Date") + ylab("Insect Abundance (Count)")+
                ggtitle(paste(paste("Insect Abundance from", startdate, "to", enddate))) +
                theme_minimal()+
                theme(legend.position = "none")

w.timeseries <- ggplot(weather, aes(x=date, y=TemperatureAvgF)) +
                geom_line() +
                xlab("Date") + ylab("Avg. Temperature in F˚")+
                ggtitle(paste(paste("Average Temperature from", startdate, "to", enddate)))+
                theme_minimal()
               
grid.newpage() 
      pushViewport(viewport(layout = grid.layout(11, 1)))
      print(map, vp = viewport(layout.pos.row = 1:7, layout.pos.col = 1))
      print(i.timeseries, vp = viewport(layout.pos.row = 8:9, layout.pos.col = 1)) 
      print(w.timeseries, vp = viewport(layout.pos.row = 10:11, layout.pos.col = 1)) 
      
    }, height = 800, width = 800 )

```
